---
title: "Assignment 3"
author: "Matt O'Reilly"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(LearnBayes)
library(dplyr)
library(ggplot2)
source("TeachBayes.r") ## put the TeachBayes.r file in the same folder as this Rmd file
```

Complete Assignment 3 using this R markdown file, building a report on each of the three statistical analyses. Submit your final report via TurnItIn to Blackboard **in Microsoft Word format**. This Assignment counts as 15% of the module grade.

## Question 1
It rains a lot in Galway. How often do you think it rains? Let's set up a Bayesian analysis to combine your belief about rain in Galway with some data. We'll treat each day as independent and suggest that $X$, the number of days it rains in Galway, is Binomially distributed, i.e. $X \sim Bin(n,\theta)$

  
(a) Select (**with justification**) an appropriate prior distribution for $\theta$, the **proportion** of days that it rains in Galway. Summarise this prior using an appropriate plot.
(**5 marks**

I specified my prior using the 0.50 and the 0.90 quantiles. From intuition I asssigned appropriate values to my p50 and p90 (where p50 represents the 0.50 quantile for $\theta$ meaning it is equally likely to be smaller or larger than p50 and p90 represents the 0.90 quantile for $\theta$ meaning it is likely with probability 0.90 to be smaller than p90). 

I selected 0.65 for my p50 because I believe it rains in galway slightly more often than 50% of the time. I then selected p90 to be 0.85 because I suspect the distribution of rainfall over a period of time to be left-skewed.

```{r}
p50 <- list(x = 0.65, p = 0.5)
p90 <- list(x = 0.85, p = 0.9)

## Find the matching beta curve
my_shapes <- beta.select(p50, p90)
my_shapes

beta_draw(my_shapes)
```

```{r}
prior_par <- c(my_shapes[1], my_shapes[2])
```

(b) Now collect data yourself on rain in Galway over the course of a few days. The total number of days on which you collect data is $n$, and the total number of days with rain is $x$. Update your prior proportion and interpret the resulting posterior probability distribution. (**10 marks**)

```{r}
data <- c(7,4)
post_par <- prior_par + data
post_par

beta_prior_post(prior_par, post_par)
```
From the updated posterior probability distribution we can see that the data has shifted slightly to the left. We can see from the shape of our posterior curve that our data is more accurate. The mode is now approximately at $\theta$ = 0.63

(c) Provide adequate summaries of the posterior including a point and interval estimate. Plot and interpret the interval estimate. (**10 marks**)
```{r}
#Point Interval Alpha/(Alpha + Beta)
point_estimate <- (11.54/(11.54+6.59))
point_estimate

beta_interval(0.95, c(my_shapes[1]+7, my_shapes[2]+4))

```
From our interval estimate we can say with 95% confidence that the probability of $\theta$ is between 0.41 and 0.835.

(d) You then look up the weather reports and find the exact same data as your own occurred last year. Update your posterior with the new data, providing summary plots and an interval estimate for the new posterior. (**6 marks**)

```{r}
prior_new <- prior_par + data
posterior_new <- prior_par + 2*data
beta_prior_post(prior_new, posterior_new) 

```

```{r}
beta_interval(0.95, posterior_new)
```
After updating our posterior with data that backs up my previous results as expected the shape of our updated posterior probability distribution is more accurate and slightly shifted to the right. The mode is still centered around a value of approximately $\theta$ = 0.6365. Looking at our updated interval estimate we can say with 95% confidence that the probability of $\theta$ is between 0.457 and 0.798.

(e) A tourist agency claims that it rains less than 20% of the time in Galway. Test this claim using your updated posterior. (**6 marks**)
```{r}
beta_area(0, 0.2,posterior_new)
```
This suggests that there is no chance that it rains less than 20% of the time in Galway. We reject the tourist agency’s claim.

(f) Simulate 10000 observations from your updated posterior (**3 marks**)
```{r}
mysims <- rbeta(10000, shape1 = posterior_new[1],shape2 = posterior_new[2])
```

(g) Using these simulated data, compute a 95% probability interval for $\theta$ and compare it to the one you found earlier. [Hint: `quantile`] (**10 marks**)

[**Q1 = 50 marks**]

```{r question1}
quantile(mysims, probs = 0.025)
quantile(mysims, probs = 0.975)
```

The 95% probability interval obtained from simulating 10,000 observations is very close to the interval we obtained earlier. We observe that the upper and lower quartiles generated from the simulated data are approximately equal to the upper and lower quartiles we found earlier. We expect as the number of simulations increases we will obtain an interval closer in value to that of the one we obtained earlier.

## Question 2

Let's try to estimate $X$ the number of emails you receive in a day. This will be a count variable since we can only get non-negative integer values.

(a) How might $X$ be described, and what distribution might best explain the data we observe? (**1 mark**)

X is a discrete outcome. Since X is a count, we can model its dependence on $\theta$ by the Poisson distribution. This would best display the data we observe.

(b) Given the model which describes these data, select (**with justification**) an appropriate prior distribution for $\theta$, the number of emails 4th year NUI, Galway students receive in a day. Summarise this prior using an appropriate plot. (**4 marks**)

I selected five emails as my mean amount of emails per day and two as my standard deviation as I belive these to be appropriate assumptions. I used my mean and standard deviation to calculate my $\beta$ and my $\alpha$.

```{r}
mean=5
sd=2
my_alpha=(mean^2)/(sd^2)
my_beta=mean/(sd^2)
x <- seq(0, 12 ,length=100)
priorx <- dgamma(x, shape=my_alpha, rate=my_beta)
plot(x, priorx, type='l',lwd=3,xlim = c(0,12),ylim = c(0,0.3),col = 'blue',
main = '', xlab = 'theta', ylab = '')

```


(c) Now collect data yourself over a few days, counting the number of emails you receive each day. Using these data, update your prior belief and form a posterior distribution for the underlying rate of emails received per day. Interpret the posterior and how it relates to your prior. (**10 marks**)

```{r}
n=10
observations=c(1,5,4,4,2,2,3,1,4,5)

post_alpha= my_alpha + sum(observations)
post_beta= my_beta + n
postx <- dgamma(x, shape= post_alpha, rate= post_beta)
plot(x, priorx, type='l',lwd=3,xlim = c(0,12),ylim = c(0,0.8), col = 'blue',
main = 'Prior', xlab = 'theta', ylab = '')
lines(x, postx,lwd=3,col='red')
legend("topright", c("Prior","Posterior"), lty = 1, lwd= 3, col = c('blue','red'))

```

From the updated posterior probability distribution we observe that the data has shifted to the left. It is evident from the shape of our posterior curve that our data is more accurate. The mode is now approximately at $\theta$ = 3.4

(d) Provide a summary of your posterior, with a point and interval estimate. Make sure to plot and interpret your interval. (**10 marks**)

```{r}
point_est= post_alpha/post_beta
point_est
## [1] 3.311111
qgamma(0.025, shape = post_alpha, rate = post_beta)
## [1] 2.334322
qgamma(0.975, shape = post_alpha, rate = post_beta)
## [1] 4.455955
plot(x, postx, type = 'l', ylim = c(0,1.2), lwd=3,col = 'red', main = 'Posterior', xlab = 'theta')
abline(v= qgamma(0.025, shape = post_alpha, rate = post_beta), lty = 'dashed', lwd = 2)
abline(v= qgamma(0.975, shape = post_alpha, rate = post_beta), lty = 'dashed', lwd = 2)

```
From the interval estimate we can conclude that the probability that $\theta$ is between 2.334322 and 4.455955 is 0.95.

(e) Another student tells you they received 10, 4, 3, 7 and 4 emails during the week. Update your posterior with the new data, providing summary plots and an interval estimate for the new posterior. (**6 marks**)
```{r}
tot_emails<-c(observations,10,4,3,7,4)

post2_alpha= my_alpha + sum(tot_emails)
post2_beta= my_beta + n
post2x <- dgamma(x, shape= post2_alpha, rate= post2_beta)
plot(x, postx, type='l',lwd=3,xlim = c(0,12),ylim = c(0,0.8), col = 'blue',
main = 'Prior', xlab = 'theta', ylab = '')
lines(x, post2x,lwd=3,col='red')
legend("topright", c("Prior","Posterior"), lty = 1, lwd= 3, col = c('blue','red'))

point2_est= post2_alpha/post2_beta
point2_est
## [1] 3.311111
qgamma(0.025, shape = post2_alpha, rate = post2_beta)
## [1] 2.334322
qgamma(0.975, shape = post2_alpha, rate = post2_beta)
## [1] 4.455955
plot(x, post2x, type = 'l', ylim = c(0,1.2), lwd=3,col = 'red', main = 'Posterior', xlab = 'theta')
abline(v= qgamma(0.025, shape = post2_alpha, rate = post2_beta), lty = 'dashed', lwd = 2)
abline(v= qgamma(0.975, shape = post2_alpha, rate = post2_beta), lty = 'dashed', lwd = 2)
```
From the new updated posterior probability distribution we observe that the data has shifted to the right. It is evident from the shape of our posterior curve that our data is more accurate. The mode is now approximately at $\theta$ = 5.8

From the interval estimate we can conclude that the probability that $\theta$ is between 4.478687 and 7.289509 is 0.95.

(f) A study has come out suggesting the total number of emails students receive is 3.2 per day. Test this claim using your posterior distribution. (**6 marks**)
```{r}
pgamma(3.2, shape = post_alpha, rate = post_beta)
```
There is a 0.4394591 probability that the total number of emails students receive is 3.2 per day on average. We fail to reject the student’s claim.

(g) Simulate 10000 observations from your updated posterior. (**3 marks**)
```{r}
mysims <- rgamma(10000, shape = post2_alpha, rate = post2_beta)
```

(h) Using these simulated data, compute a 95% probability interval for $\theta$ and compare it to the one you found earlier. [Hint: `quantile`] (**10 marks**)
```{r}
quantile(mysims, probs = c(0.025, 0.975))
```
Simliar to Q1 the 95% probability interval obtained from simulating 10000 observations is very close to the value of the interval we obtained earlier. We observe that the upper and lower quartiles generated from the simulated data are approximately equal to the upper and lower quartiles we found earlier. We expect as the number of simulations increases we will obtain an interval closer in value to that of the one we obtained earlier.

[**Q2 = 50 marks**]

## Question 3

Spending excessive time staring at a mobile phone screen (i.e. screen time) is known to have [adverse effects on children and adolescents](https://www.sciencedirect.com/science/article/pii/S001393511830015X). Less is known about phone screen time among **University students**. A person's daily screen time $X$ depends on a number of factors, including the time of day, weather, news events and sports results. Let's say that daily phone screen time is Normally distributed with mean $\theta$ and known standard deviation $\sigma$, i.e. $X\sim N(\theta,\sigma^2)$. Let's say that $\sigma$ is (magically) known to be 30 minutes.

(a) Choose (**with justification**) an appropriate prior distribution for $\theta$, the mean screen time of students. Provide summaries of this prior. (**5 marks**)
```{r}
mu0 <- 125
sigma0 <- 30
```

(b) Now collect data on daily mobile phone time, e.g. using an app for your own time or find publicly available data online if you don't use your phone! Don't worry - this doesn't need to be exact. Using the collected data, update your prior distribution and compute a posterior. Plot the **prior, _data_ and posterior** together on a single plot and interpret the relationship between them. (**15 marks**)
```{r}
n = 10
times <- c(115,122,119,102,98,119,134,145,110,98)
mean_t <- (sum(times))/n
mean_t
se = 5/sqrt(n)

post_mean = (1/(se^2) + 1/(sigma0^2))^(-1)*(mean_t/(se^2) + mu0/(sigma0^2))
post_sd = sqrt((1/(se^2) + 1/(5^2))^(-1))
df <- data.frame(Distribution = c("Prior","Data","Posterior"),
Mean = c(mu0,mean_t,post_mean),
SD = c(sigma0,se,post_sd))
df
post_mean

x <- seq(10, 145, length=10000)
priorx <- dnorm(x, mean=mu0, sd=sigma0)
datax <- dnorm(x, mean=mean_t, sd=se)
postx <- dnorm(x, mean=post_mean, sd=post_sd)
plot(x, priorx, type='l',lwd=3,xlim = c(95,145),ylim=c(0,0.3),
col = 'blue', main = '', xlab = 'theta', ylab = '')
lines(x, datax,col='black',lwd=3)
lines(x, postx,col='red',lwd=3)
legend("topleft", c("Prior","Data","Posterior"), lty = 1, lwd= 3,
col = c('blue','black','red'))


```
The posterior has shifted to the left towards the data. Our prior belief is a normal distribution $X\sim N(\theta,\sigma^2)$ with mean equal to 130 and $\sigma$ = 30. The mode is now approximately 116 minutes. We can see the distribution is more precise than both the data and prior, taking advatange of combining information from both.

(c) Comment on the shape of your data - are they likely to be come from a Normal distribution? (**2 marks**)
```{r}
qqnorm(times,main="QQ plot of normal data",pch=19)
qqline(times)

shapiro.test(times)
```
Ho: The data is normaly distributed.
Ha: The data is not normally distributed.

From the Shapiro-Wilk normality test we can see the p-value is 0.5226. This is greater than 0.05 so we fail to reject the null hypothesis.

There is evidence to suggest that the times I gathered are normally distributed.

(d) Compute a point and interval estimate for $\theta$ using your posterior and include a plot of this interval along with your interpretation. (**10 marks**)
```{r}
#Point estimate is the posterior mean
116.2244

#Interval estimate
qnorm(0.025, mean = post_mean, sd = post_sd)
qnorm(0.975, mean = post_mean, sd = post_sd)
plot(x, postx, type='l', xlim = c(105,125), ylim = c(0,0.3), lwd=3, col = 'red', main = 'Posterior',xlab = 'theta', ylab = '')
abline(v= qnorm(0.025, mean = post_mean, sd = post_sd), lty = 'dashed', lwd = 2)
abline(v= qnorm(0.975, mean = post_mean, sd = post_sd), lty = 'dashed', lwd = 2)

```
From the interval estimation we can conclude that the probability that $\theta$ lies between 113.2696 and 119.1791 is 0.95.

(e) A psychologist has published an article which estimates that students' average daily screen time is 90 minutes. Test this claim using your own posterior distribution. (**5 marks**)
```{r}
1 - pnorm(90, mean=post_mean,sd=post_sd)
```
There is a probability of 1 that the average daily screen time for a student is 30 minutes or longer. We reject the psychologists claim.

(f) Simulate 10000 values from your posterior distribution for $\theta$. (**3 marks**)
```{r}
my_sims <- rnorm(10000, mean = post_mean, sd = post_sd)
```

(g) Using the simulations, compute a 95% probability interval and compare it with the one you found earlier. [Hint: `quantile`] (**10 marks**)
```{r}
quantile(my_sims, probs = c(0.025, 0.975))
```
Like our other examples, the 95% probability interval obtained from simulating 10,000 observations is very close to the interval we obtained earlier. We observe that the upper and lower quartiles generated from the simulated data are approximately equal to the upper and lower quartiles we found earlier. We expect as the number of simulations increases we will obtain an interval closer in value to that of the one we obtained earlier.

[**Q3 = 50 marks**]

```{r  question3}


```

[**Total marks = 150**]